---
title: "P8105 Homework 1"
author: "Navya Koneripalli"
date: "2023-09-20"
output: github_document
---
# Question 1
## Setup
This chunk loads necessary packages and datasets
```{r setup}
library(tidyverse)
library(moderndive)
data("early_january_weather")
```

## Description of the data:

The "early_january_dataset" contains hourly temperature, dewpoint, humidity, wind direction, windspeed, wind gust, precipitation and visibility measurements from EWR for the first 15 days of January 2013.

### Number of rows in the dataset
```{r # of rows in the data set}
print(nrow(early_january_weather)) #printing the number of rows in "early_january_dataset"
```

There are 358 rows in the "early_january_dataset".

### Number of columns in the dataset
```{r # of columns in the data set}
print(ncol(early_january_weather)) #printing the number of columns in "early_january_dataset"
```

There are 15 columns in the "early_january_dataset".

### Mean temperature
```{r calculating the mean temperature}
mean_temp = mean(early_january_weather$temp) #creating a variable called mean_temp for the average of the hourly temperature variable in the early_january_dataset.
print(mean_temp) #printing the mean_temp variable
```

The mean hourly temperature measurement taken from EWR in the first 15 days of January 2013 is 39.58212 degrees.

## Creating a scatterplot
```{r creating a scatterplot}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point() + #created a scatterplot with time_hour (x-axis) vs temp (y-axis) where the scatterplot points are colored based on the "humid" variable
labs(x = "Hourly Timestamp for Jan 1-15, 2013", y = "Temperature (ºC)", color = "Humidity (%)") #Modified the scatterplot labels for ease of understanding

```

### Patterns in the plot:
Overall, the trend is that the temperature rose between Jan 1, 2013 and Jan 15, 2013.
Generally, the humidity percentage also increased between Jan 1, 2013 and Jan 15, 2013.
The plot also shows a daily high temperature during the daytime for each day and a drop in temperature during the nighttime.

## Saving the scatterplot
```{r saving the scatterplot}
ggsave("P8105_hw1_nk3037.pdf") #saved the plot as a .pdf file named "P8105_hw1_nk3037")
```


# Question 2

## Creating a dataframe
```{r creating a dataframe}
question2_df =
    tibble(
      vec_numeric = rnorm(10), #vector called vec_numeric that contains 10 randomly generated numbers that are normally distributed
      vec_logical = vec_numeric > 0, #vector called vec_logical that contains logical values of whether the numbers in random_sample are greater than 0
      vec_char = c("L", "M", "N", "O", "P", "Q", "R", "S", "T", "U"), #vector containing 10 characters
      vec_factor = factor(sample(1:3, 10, replace = TRUE), levels = c(1, 2, 3)) #factor vector of length 10, with 3 different factor “levels”
    )
```

## Taking the mean of each variable

### Calculating the mean of vec_numeric
```{r Calculating the mean of vec_numeric}
mean_vec_numeric = mean(pull(question2_df, vec_numeric)) #using the pull function to pull vec_numeric out of the dataframe and calculating the mean
print(mean_vec_numeric) #printing the newly created mean_vec_numeric variable
```
**It is possible to take a mean of the vec_numeric variable in the question2_df**

### Calculating the mean of vec_logical
```{r Calculating the mean of vec_logical}
mean_vec_logical = mean(pull(question2_df, vec_logical)) #using the pull function to pull vec_logical out of the dataframe and calculating the mean
print(mean_vec_logical) #printing the newly created mean_vec_logical variable
```
**It is possible to take a mean of the vec_logical variable in the question2_df**

This is because R reads the logical value TRUE as 1 and FALSE as 0.

### Calculating the mean of vec_char
```{r Calculating the mean of vec_char}
mean_vec_char = mean(pull(question2_df, vec_char)) #using the pull function to pull vec_char out of the dataframe and calculating the mean
print(mean_vec_char) #printing the newly created mean_vec_char variable
```
**It is *NOT* possible to take a mean of the vec_char variable in the question2_df**

### Calculating the mean of vec_factor
```{r Calculating the mean of vec_factor}
mean_vec_factor = mean(pull(question2_df, vec_factor)) #using the pull function to pull vec_factor out of the dataframe and calculating the mean
print(mean_vec_factor) #printing the newly created mean_vec_factor variable
```
**It is *NOT* possible to take a mean of the vec_factor variable in the question2_df**

## Converting variable types

### Applying the as.numeric function to the logical, character, and factor variables 
```{r, results ='hide', warning=FALSE}
print(as.numeric(pull(question2_df, vec_logical)))
print(as.numeric(pull(question2_df, vec_char)))
print(as.numeric(pull(question2_df, vec_factor)))
```


It is possible to find the mean of a logical variable since R reads TRUE as 1 and FALSE as 0. When as.numeric() is applied to vec_logical, the resulting vector will contain the numerical values of 0 and 1.

It is not possible to find the mean of a character variable. When as.numeric() is applied to characters, R is unable to convert the character data type into numeric data type and so the resulting vector contains only NA.

It is not possible to find the mean of a factor variable. When as.numeric() is applied to factors, R converts each level into its corresponding numeric value (level 3 is stored as "3"). The resulting vector contains numeric values (in this case the vector contains 1,2 and 3 because the factor vector was defined to have three levels)